Technique for \earlywarning\ detection of compact binary coalescence and its implications for multi-messenger astronomy

### 2011-06-10 method.tex
Reorganize section 2.2 to make a less grumpy grad students even happier.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1010 ccdd05d5-56c1-461a-bf98-85f6876523b4

Here we explore a method for reducing the computational cost of a \TD\ search
for compact binary coalescence.  We will give a truly zero-latency algorithm
that competes in terms of floating point operations per second with the
conventional \FD\ method, which requires a significant latency in order to be
computationally cheap. Our method, called \lloid{} (a backronym for
Low Latency Online Inspiral Detection),
involves two transformations of the template waveforms that produce a set of
orthogonal filters with far fewer coefficients than the original templates.

The first transformation is to chop the \TD{} templates up into \emph{time
slices}.  Since each template slice is disjoint in time, the resulting set for
a single template is orthogonal.  Given the chirp-like structure of the
templates, the early time slices have significantly lower bandwidth and can be
safely down-sampled.  The down-sampling reduces the number of data samples by a
factor of $\sim$100 from the early part of the waveform and allows the filters
to be evaluated at about $\sim$$1/100^\mathrm{th}$ the sample rate. This
amounts to more than a factor of 10000
reduction in the floating-point operations per second required to filter the
early part of the waveform.  However, the resulting filters are still not
orthogonal across the mass parameter space, and are in fact highly redundant.
We use the \SVD{} to produce an orthogonal filter set
from the time-sliced templates~\citep{Cannon:2010p10398}.  We find that this
reduces the number of sample points of the filters by another factor of
$\sim$100.  The combined methods reduce the number of floating point operations
to the level where they are competitive with the conventional high-latency \FD\
matched filter approach.  In the remainder of this section we describe the
\lloid\ algorithm in detail and provide some basic computational cost scaling.

>>>>

Here we describe a method for reducing the computational cost of a \TD\ search
for compact binary coalescence.  We give a truly zero-latency algorithm
that competes in terms of floating point operations per second with the
conventional \FD\ method, which requires a significant latency in order to be
computationally cheap. Our method, called \lloid{} (a backronym for
Low Latency Online Inspiral Detection),
involves two transformations of the template waveforms that produce a
smaller set of \emph{orthogonal} filters with far fewer coefficients than
 the original templates.

The first transformation is to chop the \TD{} templates up into \emph{time
slices}.  Since each template slice is disjoint in time, the resulting set for
a single template is orthogonal.  Given the chirp-like structure of the
templates, the early time slices have significantly lower bandwidth and can be
safely down-sampled.  The down-sampling reduces the number of data samples by a
factor of $\sim$100 from the early part of the waveform and allows the filters
to be evaluated at about $\sim$$1/100^\mathrm{th}$ the sample rate. Since the
computational cost of \TD\ filtering is quadratic in the filter length, this
mounts to more than a factor of $10^4$ reduction in the floating-point operations
 per second required to filter the early part of the waveform.

However, the resulting filters are still not
orthogonal across the mass parameter space, and are in fact highly redundant.
We use the \SVD{} to approximate the template bank by a set of orthogonal filters
~\citep{Cannon:2010p10398}.  We find that this approximation
reduces the number of filters needed by another factor of
$\sim$100.  The combined methods reduce the number of floating point operations
to the level where they are competitive with the conventional high-latency \FD\
matched filter approach.  In the remainder of this section we describe the
\lloid\ algorithm in detail and provide some basic computational cost scaling.
